---
title: "EDS 230 Homework 6: Using Sobol with an ODE"
format: pdf
author: Leela Dixit, Jay Kim, Richard Montes Lemus
---


```{r setup, include=FALSE}
library(tidyverse)
library(deSolve)
library(here)
library(sensitivity)
```

### Forest Growth Model Implementation
```{r}
# source population growth model
source(here("R/dforestgrowth.R"))
```

### Running the Model with Initial Parameters
```{r}
# initialize initial conditions
Cinitial <- 10
tm <- seq(from = 1, to = 300)
parms = list(K = 250, r = 0.01, g = 2, thresh = 50)

ode <- ode(Cinitial, tm, dforestgrowth, parms)

colnames(ode) <- c("time", "forest_size")
head(ode)
tail(ode)
```
### Visualizing Forest Growth Trajectory
```{r}
ggplot(as.data.frame(ode), aes(x = time, y = forest_size)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Time series of forest growth with canopy\nclosure threshold at 50 kgC",
       x = "Years",
       y = "Forest Size (Carbon)")
```

### Run a sobol global sensitivity analysis with parameter sampling

We are interested in estimating the **maximum forest size** varies with parameter variation over a time period of 300 years. We will assume all parameters are normally distributed.

```{r sen, message=FALSE}
# Assume initial forest size is 10 kg/C
Cinitial <- 10

# set the number of parameter samples you have as 2000
np <- 2000

# generate parameters assuming normal distribution
K <- rnorm(mean = 250, sd = 25, n = np)
r <- rnorm(mean = 0.01, sd = 0.001, n = np)
g <- rnorm(mean = 2, sd = 0.2, n = np)
thresh <- rnorm(mean = 50, sd = 5, n = np)

# join into a data frame
X1 <- cbind.data.frame(r = r, K = K, g = g, thresh = thresh)

# repeat to get our second set of samples
K <- rnorm(mean = 250, sd = 25, n = np)
r <- rnorm(mean = 0.01, sd = 0.001, n = np)
g <- rnorm(mean = 2, sd = 0.2, n = np)
thresh <- rnorm(mean = 50, sd = 5, n = np)
X2 <- cbind.data.frame(r = r, K = K, g = g, thresh = thresh)

# fix any negative values and they are not meaningful (using pmax and map_df)
X1 <- X1 %>% 
  map_df(pmax, 0.0)
X2 <- X2 %>% 
  map_df(pmax, 0.0)

# create our sobol object and get sets of
#parameters for running the model, use sobolSalt
sens_C <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# take a look at parameter sets
str(sens_C)

# lets add names
colnames(sens_C$X) <- c("r", "K", "g", "thresh")

head(sens_C$X)
```


```{r}
# run for 300 days
simtimes = seq(from = 1, to = 300)

# use the first parameter set (or you could just pick a test value), 
# not the list because ode wants parameters as a list, so name each
parmtest = list(r = sens_C$X[1, "r"], 
                K = sens_C$X[1, "K"],
                g = sens_C$X[1, "g"],
                thresh = sens_C$X[1, "thresh"])
# run ode
result <-  ode(y = Cinitial, 
               times = simtimes, 
               func = dforestgrowth, 
               parms = parmtest)
# turn ode results into a labelled data frame 
result <- data.frame(time = result[,1], C = result[,2])

# create function to pull out metrics we want 
compute_metrics <- function(result) {
  maxsize <- max(result$C)
  return(maxsize)
}
# run metrics on results, lets use a threshold of 100
compute_metrics(result)

# wrap ode solver and computing metrics 
p_wrapper <- function(r, K, g, 
                      thresh, 
                      Cinitial, 
                      simtimes, 
                      odefunc, 
                      metricfunc) {
  
  # create parameter list
  parms = list(r=r, K=K, g=g, thresh=thresh)
  # run ode
  result = ode(y = Cinitial, 
               times = simtimes, 
               func = odefunc, 
               parms = parms)
  # results as a data frame
  result = data.frame(time = result[,1], C = result[,2])
  # get and compute metrics
  metric = metricfunc(result)
  return(metric)
}

# test with specific r and K values (r=0.01, K=150) and Cinitial = 10 
# just do 10 time steps
p_wrapper(r = 0.01, K = 150, g = 2, thresh = 50,
          Cinitial = 10, 
          simtimes = seq(from = 1, to = 10),
          odefunc = dforestgrowth, 
          metricfunc = compute_metrics)

# run the wrapper on full sobol samples
allresults <- as.data.frame(sens_C$X) %>% 
  pmap(p_wrapper, 
       Cinitial = Cinitial, 
       simtimes = simtimes, 
       odefunc = dforestgrowth, 
       metricfunc = compute_metrics)

# pull out data from results list
allres <- as.data.frame(allresults) %>% 
  pivot_longer(cols = everything(),
               names_to = "metric",
               values_to = "maxsize")
```

### Graph sobol results
```{r}
ggplot(allres, aes(x = maxsize)) +
  geom_boxplot()
```
### Record sobol indices
```{r}
sens_C_maxsize <- sensitivity::tell(sens_C, allres$maxsize)

rownames(sens_C_maxsize$S) <- c("r", "K", "g", "thresh")
sens_C_maxsize$S

rownames(sens_C_maxsize$T) <- c("r", "K", "g", "thresh")
sens_C_maxsize$T
```

### Conclusion: 

In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about how what parameters climate change might influence).

Our sensitivity analysis shows that maximum forest size is most influenced by the carrying capacity (K) and early growth rate (r), these explain a majority of the variation in outcomes. Post-canopy growth rate (g) contributed moderately, while canopy closure threshold had little influence on how large the forest ultimately gets. This has large implications for climate change, as both K and r could be reduced by increasing extreme weather events - leading to a smaller forests. 
